---
title: "Cocoa Price Forecasting in Ghana: Analyzing Market and Climate Impacts"
format: pdf
editor: visual
---

```{r, include=FALSE}
install.packages("astsa")
install.packages("forecast")
install.packages("tseries")
install.packages("quadprog")
install.packages('vars')
install.packages(c( "dynlm", "lmtest", "AER", "ARDL"))
```

```{r, include=FALSE}
library(astsa)
library(forecast)
library(ggplot2)
library(tidyverse)
library(dplyr)
library(caret)
library(quadprog)
library(tseries)
library(tidyr)
library(urca)
library(vars)
library(dynlm)   # Dynamic regression models
library(lmtest)  # Model testing
library(AER)     # Additional econometric functions
library(ARDL)    # ARDL model estimation
install.packages("strucchange")  # Install if you haven't already
library(strucchange)
```

```{r Data Cleaning, include=FALSE}

yearly_data <- read.csv("yearly_final - final.csv")
ghana_data <- read.csv("ghana_final - final.csv")
cameroon_data <- read.csv("cameroon_final - final.csv")
ivorycoast_data <- read.csv("ivorycoast_final.csv")
nigeria_data <- read.csv("nigera_final - final.csv")
icco <- read.csv("forecasting_icco_final.csv")


icco$daily_price <- as.numeric(gsub(",", "", icco$daily_price))
icco$all_date <- as.Date(icco$all_date)
ghana_data$ghana_date <- as.Date(ghana_data$ghana_date)
yearly_data$year <- as.double(yearly_data$year)
cameroon_data$cameroon_date <- as.Date(cameroon_data$cameroon_date)
ivorycoast_data$ivorycoast_date <- as.Date(ivorycoast_data$ivorycoast_date)
nigeria_data$nigeria_date <- as.Date(nigeria_data$nigeria_date)

icco <- icco[icco$all_date >= as.Date("1994-10-10"), ] 

head(icco)
head(ghana_data)
head(yearly_data)
head(cameroon_data)
head(ivorycoast_data)
head(nigeria_data)


```

```{r Ghana Imputation, include=FALSE}

ghana_data$ghana_ex_rate[is.na(ghana_data$ghana_ex_rate)] <- mean(ghana_data$ghana_ex_rate, na.rm = TRUE)

ghana_data$ghana_avg_temp[is.na(ghana_data$ghana_avg_temp)] <- mean(ghana_data$ghana_avg_temp, na.rm = TRUE)

summary(ghana_data)
sum(is.na(ghana_data))

```

```{r Cameroon Imputation, include=FALSE}

cameroon_data$cameroon_ex_rate[is.na(cameroon_data$cameroon_ex_rate)] <- mean(cameroon_data$cameroon_ex_rate, na.rm = TRUE)

cameroon_data$cameroon_avg_temp[is.na(cameroon_data$cameroon_avg_temp)] <- mean(ghana_data$ghana_avg_temp, na.rm = TRUE)

summary(cameroon_data)
```

```{r Ivory Coast Imputation, include=FALSE}

ivorycoast_data$ivorycoast_ex_rate[is.na(ivorycoast_data$ivorycoast_ex_rate)] <- mean(ivorycoast_data$ivorycoast_ex_rate, na.rm = TRUE)

ivorycoast_data$ivorycoast_avg_temp[is.na(ivorycoast_data$ivorycoast_avg_temp)] <- mean(ivorycoast_data$ivorycoast_avg_temp, na.rm = TRUE)

summary(ivorycoast_data)
```

```{r Nigeria Imputation, include=FALSE}

nigeria_data$nigeria_ex_rate[is.na(nigeria_data$nigeria_ex_rate)] <- mean(nigeria_data$nigeria_ex_rate, na.rm = TRUE)

nigeria_data$nigeria_avg_temp[is.na(nigeria_data$nigeria_avg_temp)] <- mean(nigeria_data$nigeria_avg_temp, na.rm = TRUE)

summary(nigeria_data)
```

```{r ICCO Imputation, include=FALSE}

# Replace NA with the previous non-NA value
icco <- icco %>%
  fill(daily_price, .direction = "down")

summary(icco)

```

# Introduction

# Literature Review

# Methodology

# Data

# Forecasting and Results

```{r}




```

```{=tex}
\begin{align*}
Y_t &= 1.486 \times 10^3 + 1.199 Y_{t-1} + 1.249 Y_{t-2} + 8.837 \times 10^{-2} Y_{t-3} \\
&\quad + 1.150 Y_{t-4} + 7.431 \times 10^1 Y_{t-5} \\
&\quad + (3.543 \times 10^5) X_{1,t} + (2.050 \times 10^5) X_{1,t-1} - (7.765 \times 10^3) X_{1,t-2} \\
&\quad - (6.272 \times 10^5) X_{1,t-3} - (1.482 \times 10^5) X_{1,t-4} - (1.339 \times 10^5) X_{1,t-5} \\
&\quad - (2.995 \times 10^1) X_{2,t} + 1.569 X_{2,t-1} - 1.579 X_{2,t-2} \\
&\quad - 6.612 X_{2,t-3} + 5.887 X_{2,t-4} - 3.244 X_{2,t-5} \\
&\quad - (2.766 \times 10^2) X_{3,t} + 1.871 X_{3,t-1} - 1.862 X_{3,t-2} \\
&\quad + 1.693 X_{3,t-3} - 1.964 X_{3,t-4} - 1.290 X_{3,t-5} \\
&\quad - (9.584 \times 10^2) X_{4,t} + 8.297 X_{4,t-1} + 2.917 X_{4,t-2} \\
&\quad + 7.165 X_{4,t-3} - 3.507 X_{4,t-4} + 1.150 T_t + \epsilon_t
\end{align*}
```
```{=tex}
\begin{align*}
Y_t &= \beta_0 + \beta_1 Y_{t-1} + \beta_2 Y_{t-2} + \beta_3 Y_{t-3} + \beta_4 Y_{t-4} + \beta_5 Y_{t-5} \\
&\quad + \gamma_0 X_{1,t} + \gamma_1 X_{1,t-1} + \gamma_2 X_{1,t-2} + \gamma_3 X_{1,t-3} + \gamma_4 X_{1,t-4} + \gamma_5 X_{1,t-5} \\
&\quad + \delta_0 X_{2,t} + \delta_1 X_{2,t-1} + \delta_2 X_{2,t-2} + \delta_3 X_{2,t-3} + \delta_4 X_{2,t-4} + \delta_5 X_{2,t-5} \\
&\quad + \theta_0 X_{3,t} + \theta_1 X_{3,t-1} + \theta_2 X_{3,t-2} + \theta_3 X_{3,t-3} + \theta_4 X_{3,t-4} + \theta_5 X_{3,t-5} \\
&\quad + \lambda_0 X_{4,t} + \lambda_1 X_{4,t-1} + \lambda_2 X_{4,t-2} + \lambda_3 X_{4,t-3} + \lambda_4 X_{4,t-4} + \lambda_5 X_{4,t-5} \\
&\quad + \tau_0 T_t + \tau_1 T_{t-1} + \tau_2 T_{t-2} + \tau_3 T_{t-3} + \tau_4 T_{t-4} + \tau_5 T_{t-5} + \epsilon_t
\end{align*}
```
Variable Definitions

```{=tex}
\begin{itemize}
    \item \( Y_t \) = Daily price
    \item \( X_1 \) = Cameroon exchange rate
    \item \( X_2 \) = Cameroon average temperature
    \item \( X_3 \) = Ghana exchange rate
    \item \( X_4 \) = Ghana average temperature
    \item \( T_t \) = Trend variable
    \item \( \epsilon_t \) = Error term
\end{itemize}
```
```{r}
#giving a standard name to the dates column

colnames(nigeria_data)[colnames(nigeria_data)=='nigeria_date'] <- 'Date'
colnames(cameroon_data)[colnames(cameroon_data)=='cameroon_date'] <- 'Date'
colnames(ghana_data)[colnames(ghana_data)=='ghana_date'] <- 'Date'
colnames(ivorycoast_data)[colnames(ivorycoast_data)=='ivorycoast_date'] <- 'Date'
colnames(icco)[colnames(icco)=='all_date'] <- 'Date'




```

```{r}
tail(icco)

```

```{r}
#merging dataframes

dfs <- list(icco,nigeria_data,ghana_data,ivorycoast_data,cameroon_data)

merged_df <- Reduce(function(x,y) merge(x,y, by='Date', all=TRUE),dfs)

merged_df <- na.omit(merged_df)

tail(merged_df)
```

```{r}
#grouping values by month

merged_df <- merged_df %>% mutate(YearMonth = format(Date, "%Y-%m")) %>%
  group_by(YearMonth) %>% summarise(daily_price = mean(daily_price, na.rm = TRUE), nigeria_ex_rate = mean(nigeria_ex_rate), nigeria_avg_temp = mean(nigeria_avg_temp), ghana_ex_rate = mean(ghana_ex_rate), ghana_avg_temp = mean(ghana_avg_temp), cameroon_ex_rate = mean(cameroon_ex_rate), cameroon_avg_temp = mean(cameroon_avg_temp), ivorycoast_ex_rate = mean(ivorycoast_ex_rate), ivorycoast_avg_temp = mean(ivorycoast_avg_temp))

head(merged_df)

merged_df

```

```{r}

library(zoo)

merged_df$YearMonth <- as.yearmon(merged_df$YearMonth, format = "%Y-%m")
merged_df <- na.omit(merged_df)

ts_data <- ts(merged_df[, -1], start = c(2007, 7), frequency = 12) 


```

```{r}
plot(ts_data[,1],type = "l", col = "blue", 
     main = "Time Series Plot for daily price", 
     xlab = "Time", ylab = "monthly price")

plot(ts_data[,2],type = "l", col = "blue", 
     main = "Time Series Plot for nigerian exchange rate", 
     xlab = "Time", ylab = "monthly Nigerian exchange rate")

plot(ts_data[,3],type = "l", col = "blue", 
     main = "Time Series Plot for nigerian average temperature", 
     xlab = "Time", ylab = "monthy average temperature")

plot(ts_data[,4],type = "l", col = "blue", 
     main = "Time Series Plot for ghana exchange rate", 
     xlab = "Time", ylab = "monthly ghana exchange rate")

plot(ts_data[,5],type = "l", col = "blue", 
     main = "Time Series Plot for ghana average temperature", 
     xlab = "Time", ylab = "monthy average temperature")

plot(ts_data[,6],type = "l", col = "blue", 
     main = "Time Series Plot for Cameroon exchange rate", 
     xlab = "Time", ylab = "monthly Cameroon exchange rate")

plot(ts_data[,7],type = "l", col = "blue", 
     main = "Time Series Plot for Cameroon average temperature", 
     xlab = "Time", ylab = "monthy average temperature")

plot(ts_data[,8],type = "l", col = "blue", 
     main = "Time Series Plot for ivorycoast exchange rate", 
     xlab = "Time", ylab = "monthly ivorycoast exchange rate")

plot(ts_data[,9],type = "l", col = "blue", 
     main = "Time Series Plot for ivorycoast average temperature", 
     xlab = "Time", ylab = "monthy average temperature")
```

```{r}

#testing for co-integration

# Define response variable (Y) and predictors (X1, X2, ..., X8)
response_var <- "daily_price"  # Replace with the actual response variable name
predictor_vars <- c("nigeria_ex_rate", "nigeria_avg_temp", 
                    "ghana_ex_rate", "ghana_avg_temp", "cameroon_ex_rate", "cameroon_avg_temp", "ivorycoast_ex_rate", "ivorycoast_avg_temp")  # Replace with actual predictor names

# Create a dataframe to store results
adf_results <- data.frame(Predictor = character(), 
                          ADF_Stat = numeric(), 
                          P_Value = numeric(), 
                          Critical_1pct = numeric(), 
                          Critical_5pct = numeric(), 
                          Critical_10pct = numeric(), 
                          stringsAsFactors = FALSE)

for (predictor in predictor_vars) {
  
  # Run linear regression: Response ~ Predictor
  model <- lm(as.formula(paste(response_var, "~", predictor)), data = ts_data)
  
  # Extract residuals
  residuals_model <- residuals(model)
  
  # Perform ADF test on residuals
  adf_test <- ur.df(residuals_model, type = "none", lags = 5)
  
  # Extract results
  adf_stat <- adf_test@teststat[1]  # ADF test statistic
  critical_values <- adf_test@cval[1, ]  # Critical values
  
  # Store results in dataframe
  adf_results <- rbind(adf_results, data.frame(
    Predictor = predictor,
    ADF_Stat = adf_stat,
    Critical_1pct = critical_values["1pct"],
    Critical_5pct = critical_values["5pct"],
    Critical_10pct = critical_values["10pct"]
  ))
}

# Print the results
print(adf_results)
```

We can see clearly that co-integration exists for all our predictors in the time series. This is because we reject the null hypothesis that the residuals are non-stationary at all confidence intervals for all the series. In other words, all residuals are stationary i.e all country-based

exchange rates and average temperatures co-integrate with Cocoa commodity futures prices.

This warrants the use of an error-correction model on top of an ARDL model.

```{r}
# testing for stationarity

series_names <- colnames(ts_data)  # Extract column names

# Function to perform ADF test and extract test stat
adf_test_results <- function(series, name) {
  test <- ur.df(series, type = "none", lags = 5)  # ADF Test
  # Extract test statistic
  test_stat <- summary(test)@teststat[1]  
  
  # Extract critical values
  critical_values <- test@cval[1, ]  # 1% , 5% , 10% levels
  
  return(data.frame(Series = name, 
                    ADF_Stat = test_stat, 
                    Critical_1pct = critical_values["1pct"], 
                    Critical_5pct = critical_values["5pct"], 
                    Critical_10pct = critical_values["10pct"]))
  
  
}

# Apply ADF test to each column in merged_df
results <- do.call(rbind, lapply(series_names, function(col) adf_test_results(ts_data[, col], col)))

# Print results
print(results)



```

Based on our results, we reject the null hypothesis that Ghana's exchange rate is non-stationary at all confidence intervals which means its stationary.

All the other series are non-stationary at all confidence intervals. In other words, we fail to reject

the null hypothesis that the series are non-stationary at all confidence intervals

```{r}
compute_diffs <- function(series_name, data) {
  order <- ndiffs(data[,series_name], test='adf')  # Compute differencing order
  return(data.frame(Series = series_name, Differencing_Order = order))
}

# Apply the function to each column in the dataframe
results_diff <- bind_rows(lapply(colnames(ts_data), compute_diffs, data = ts_data))


# Print the table
print(results_diff)



```

```         
```

All the average temperatures seem to have an order of zero. However, they may still

be trend stationary. We know from before that Ghana's exchange rate is already stationary.

Thus, since our data is either I(0) or I(1) we can use the ARDL model.

```{r}


```

Since the average temperatures for cameroon, Nigeria, ivorycoast and Ghana do not seem to be

difference stationary, we will test for trend-stationarity and structural break analysis.

```{r}

#trend stationarity and structural break tests

#cameroon avg temp
adf_test_trend_cam <- ur.df(ts_data[,'cameroon_avg_temp'], type = "trend", lags = 1)
summary(adf_test_trend_cam)

#summary(ur.za(merged_df$cameroon_avg_temp, model = "both"))

#ghana avg temp
adf_test_trend_ghana <- ur.df(ts_data[,'ghana_avg_temp'], type = "trend", lags = 1)
summary(adf_test_trend_ghana)

#summary(ur.za(merged_df$ghana_avg_temp, model = "both"))

#nigeria avg temp
adf_test_trend_nig <- ur.df(ts_data[,'nigeria_avg_temp'], type = "trend", lags = 1)
summary(adf_test_trend_nig)

summary(ur.za(merged_df$daily_price, model = "both"))



```

All seem to be trend stationary which means we will have to add a trend term when incorporating the ARDL model.

```{r}

# Building the optimal ARDL model

trend <- seq_along(ts_data[, 1])  # Create a time series of increasing numbers (1, 2, 3, ...)


# Add the trend variable to the data frame
merged_df$trend <- trend




set.seed(123)

train_data <- merged_df[0:205,]
test_data <- merged_df[201:nrow(merged_df),] 

head(test_data)


optimal_model_train <- auto_ardl(daily_price ~ cameroon_ex_rate + cameroon_avg_temp + ghana_ex_rate + ghana_avg_temp + trend, data = train_data[, -1], max_order = 5)

optimal_model_train$best_order

```

Typically, for the **Bounds Test**, the `case` can be one of the following:

-   **1**: No intercept, no trend

-   **2**: Intercept, no trend

-   **3**: Intercept, with trend

-   **4**: No intercept, with trend

```{r}



# Ensure trend is numeric
merged_df$trend <- as.numeric(1:nrow(merged_df))

# make columns numeric data type
merged_df$daily_price <- as.numeric(merged_df$daily_price)
merged_df$nigeria_ex_rate <- as.numeric(merged_df$nigeria_ex_rate)
merged_df$ghana_avg_temp <- as.numeric(merged_df$ghana_avg_temp)
merged_df$nigeria_avg_temp <- as.numeric(merged_df$nigeria_avg_temp)
merged_df$ghana_ex_rate <- as.numeric(merged_df$ghana_ex_rate)
merged_df$cameroon_avg_temp <- as.numeric(merged_df$cameroon_avg_temp)
merged_df$cameroon_ex_rate <- as.numeric(merged_df$cameroon_ex_rate)
merged_df$ivorycoast_ex_rate <- as.numeric(merged_df$ivorycoast_ex_rate)
merged_df$ivorycoast_avg_temp <- as.numeric(merged_df$ivorycoast_avg_temp)

merged_df <- na.omit(merged_df)



best_model_train <- ardl(daily_price ~ cameroon_ex_rate + cameroon_avg_temp + ghana_ex_rate + ghana_avg_temp + trend,
  data = train_data[, -1],
 order = c(5,5,5,5,5,5))


# Bounds Test for Cointegration with multiple independent variables
uecm_model <- uecm(best_model_train, data=train_data[, -1])
#bounds_f_test(best_model_1, case = 3)

summary(best_model_train)

checkresiduals(best_model_train)


```

residuals are certainly not normally distributed.

```{r}
# Correlation matrix

cor(merged_df[, c("nigeria_ex_rate", "cameroon_ex_rate", "ghana_ex_rate", "ivorycoast_ex_rate")])

cor(merged_df[, c("nigeria_avg_temp", "cameroon_avg_temp", "ghana_avg_temp", "ivorycoast_avg_temp")])





```

```{r}

```

```{r}


```

```{r}
#table of results

get_model_metrics <- function(model) {
  residuals <- residuals(model)
  
  # Perform Tests
  ljung <- Box.test(residuals, lag = 10, type = "Ljung-Box")$p.value
  bp <- bptest(model)$p.value
  shapiro <- shapiro.test(residuals)$p.value
  aic_val <- AIC(model)
  bic_val <- BIC(model)
  
  # Return Results
  return(c(Ljung_Box = ljung, BP_Test = bp, Shapiro_Test = shapiro, 
           AIC = aic_val, BIC = bic_val))
}

model_metrics <- rbind(
  Model_without_2324 = get_model_metrics(best_model_train),model_eucm = get_model_metrics(uecm_model)
  
)

# Convert to Data Frame for Better Viewing
model_metrics_df <- as.data.frame(model_metrics)
print(model_metrics_df)
```

```{r}



```

```{r}
#calculting training and testing RMSE

L <- function(x, k) dplyr::lag(x, k)
#max_lag <- max(c(2, 5, 0, 0, 0, 2))  # Get the highest lag in your model

# Ensure `test_set` has enough historical data
#test_data <- bind_rows(tail(train_data, max_lag), test_data)

predictions_test <- predict(best_model_train, newdata=test_data,se.fit=TRUE)




actual_values_test <- test_data$daily_price

actual_values_train <- train_data$daily_price



predictions_test_final <- predictions_test$fit



#training predictions
predictions_train <- predict(best_model_train, newdata=train_data,se.fit=TRUE)


# Now you can calculate performance metrics for the trimmed test data
rmse_test <- sqrt(mean((actual_values_test - predictions_test_final)^2, na.rm = TRUE))
mae <- mean(abs(actual_values_test - predictions_test_final), na.rm = TRUE)

rmse_train <- sqrt(mean((actual_values_train - predictions_train$fit)^2, na.rm = TRUE))



cat("Testing Root Mean Squared Error (MSE):", rmse_test, "\n")
cat("Mean Absolute Error (MAE):", mae, "\n")
cat("Training Root Mean Squared Error (MSE):", rmse_train, "\n")

#max(merged_df_break$daily_price)

```

```{r}
# Load ggplot2 package
library(ggplot2)

# Assuming your actual and forecasted values are in vectors: actual_prices and forecasted_prices
# And you have a time variable or index (e.g., time_index)
forecast_se <- predictions_test$se.fit
data <- data.frame(
  Time = 1:(length(actual_values_test)-5),
  Actual = actual_values_test[6:length(actual_values_test)],
  Forecasted = predictions_test_final[6:length(actual_values_test)],
  Lower_Bound = predictions_test_final[6:length(actual_values_test)] - 1.96 * forecast_se[6:length(actual_values_test)],  # 95% Lower CI
  Upper_Bound = predictions_test_final[6:length(actual_values_test)] + 1.96 * forecast_se[6:length(actual_values_test)]   # 95% Upper CI
)


# Create the line plot
ggplot(data, aes(x = Time)) +
  geom_line(aes(y = Actual, color = "Actual"), size = 1) +    # Actual values line
  geom_line(aes(y = Forecasted, color = "Forecasted"), size = 1, linetype = "dashed") +  geom_ribbon(aes(ymin = Lower_Bound, ymax = Upper_Bound), fill = "red", alpha = 0.2) + # Forecasted values line
  labs(title = "ARDL Forecasted vs Actual Prices (testing data)", 
       x = "Time (August 2024 - November 2024)", 
       y = "Cocoa Futures Price") +
  scale_color_manual(name = "Legend", values = c("Actual" = "blue", "Forecasted" = "red")) +
  theme_minimal()



```

```{r}
# forecasting
# Forecast next 8 periods (Dec 2024 - July 2025)

#SARIMA for cameroon temperature

install.packages('smooth')
library(smooth)

ts_data <- ts_data[0:205,]

# SARIMA for cameroon temperature
sarima_cameroon_temp <- auto.ssarima(ts_data[,'cameroon_avg_temp'])

summary(sarima_cameroon_temp)
  
#auto.arima(merged_df$cameroon_avg_temp, seasonal=TRUE,stepwise=FALSE, approximation=FALSE, trace=FALSE)

checkresiduals(sarima_cameroon_temp)

Box.test(sarima_cameroon_temp$residuals, lag = 10, type = "Ljung-Box")$p.value


# SARIMA for cameroon exchange rate
sarima_cameroon_ex <- auto.ssarima(ts_data[,'cameroon_ex_rate'])

summary(sarima_cameroon_ex)
  
#auto.arima(merged_df$cameroon_avg_temp, seasonal=TRUE,stepwise=FALSE, approximation=FALSE, trace=FALSE)

checkresiduals(sarima_cameroon_ex)

Box.test(sarima_cameroon_ex$residuals, lag = 10, type = "Ljung-Box")$p.value


#SARIMA for ghana temperature
sarima_ghana_temp <- auto.ssarima(ts_data[,'ghana_avg_temp'])

summary(sarima_ghana_temp)
  
#auto.arima(merged_df$cameroon_avg_temp, seasonal=TRUE,stepwise=FALSE, approximation=FALSE, trace=FALSE)

checkresiduals(sarima_ghana_temp)

Box.test(sarima_ghana_temp$residuals, lag = 10, type = "Ljung-Box")$p.value



# SARIMA for ghana exchange rate
sarima_ghana_ex <- auto.ssarima(ts_data[,'ghana_ex_rate'])

summary(sarima_ghana_ex)
  
#auto.arima(merged_df$cameroon_avg_temp, seasonal=TRUE,stepwise=FALSE, approximation=FALSE, trace=FALSE)

checkresiduals(sarima_ghana_ex)

Box.test(sarima_ghana_ex$residuals, lag = 10, type = "Ljung-Box")$p.value
```

```{r}


```

```{r}



```

```{r}
training_data <- train_data
# Generate future dates from July 2024 to November 2024 for historical data
dates_historical <- seq(from = as.yearmon("2024-07", "%Y-%m"), 
                        to = as.yearmon("2024-11", "%Y-%m"), 
                        by = 1/12)

# Generate future dates from December 2024 to March 2025 for forecast
dates_future <- seq(from = as.yearmon("2024-12", "%Y-%m"), 
                    to = as.yearmon("2025-07", "%Y-%m"), 
                    by = 1/12)


# Combine both historical and forecast dates
dates_combined <- c(dates_historical, dates_future)

# Forecast each predictor using SARIMA models for the forecast period (December 2024 to March 2025)
cameroon_temp_forecast <- predict(sarima_cameroon_temp, n.ahead = length(dates_future))$mean[1:length(dates_future)]
cameroon_ex_forecast <- predict(sarima_cameroon_ex, n.ahead = length(dates_future))$mean[1:length(dates_future)]
ghana_temp_forecast <- predict(sarima_ghana_temp, n.ahead = length(dates_future))$mean[1:length(dates_future)]
ghana_ex_forecast <- predict(sarima_ghana_ex, n.ahead = length(dates_future))$mean[1:length(dates_future)]

# Initialize forecast dataframe for both historical and forecast periods
forecast_df <- data.frame(
  Yearmon = dates_combined,
  cameroon_avg_temp = c(tail(training_data$cameroon_avg_temp, 5), cameroon_temp_forecast),
  cameroon_ex_rate = c(tail(training_data$cameroon_ex_rate, 5), cameroon_ex_forecast),
  ghana_avg_temp = c(tail(training_data$ghana_avg_temp, 5), ghana_temp_forecast),
  ghana_ex_rate = c(tail(training_data$ghana_ex_rate, 5), ghana_ex_forecast),
  trend = c(tail(training_data$trend, 5), seq(length(dates_future)) + max(training_data$trend)),
  daily_price = c(tail(training_data$daily_price, 5), rep(NA, length(dates_future)))  # NA for future daily price
)

for (t in (length(dates_historical) + 1):nrow(forecast_df)) {
  # Predict the daily_price for the current row (t) using the ARDL model
  forecast_df$daily_price[t] <- predict(best_model_train, newdata = forecast_df[1:t,])[t]
}


forecast_df
#predict(best_model_1, newdata = test_data[1:7,])[7]
#test_data

# Print the forecasted dataframe


```

```{r}

library(dplyr)

# Rename forecast_df column to match training_data
forecast_df <- forecast_df %>% rename(YearMonth = Yearmon)

combined_data <- bind_rows(training_data %>% dplyr::select(YearMonth,daily_price) %>% mutate(type = "Historical"), forecast_df %>% filter(!is.na(daily_price)) %>% dplyr::select(YearMonth, daily_price) %>% mutate(type = "Forecast"))

combined_data$Date <- as.Date(as.yearmon(combined_data$YearMonth))

ggplot(combined_data, aes(x = Date, y = daily_price, color = type)) +
  geom_line(size = 1) +
  scale_color_manual(values = c("Historical" = "black", "Forecast" = "red")) +
  labs(title = "ARDL cocoa Price with Forecast",
       x = "Date",
       y = "Monthly price",
       color = "Data Type") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
```

```         
```

# Discussion and Conclusion

# Appendix

# References
